---
title: "Nonbreeding assignment using WGSassign"
author: "Matt DeSaix"
date: '2022-10-05'
output: html_document
---

I've identified snps associated with population structure and avoided ascertainment bias using a training and testing data set. Assignment accuracy is greater than 90% for the smaller SNP sets (~450 - 8000) for the K=5 population groups. Now to check assignment of the nonbreeding individuals - what's migratory connectivity going to look like??


```{r}
library(tidyverse)
```

```{r}
amre.bams <- read_table("../data/samples/lists/downsampling/amre-all-norelate_nobadPA-down2.0x-bam-list.txt",
                        col_names = c("full_path")) %>%
  mutate(Bam = gsub(pattern = "/home/mgdesaix@colostate.edu/scratch/AMRE/all-samples/bamfiles/xfull/",
                    replacement = "",
                    x = full_path)) %>%
  mutate(Bam = gsub(pattern = "/home/mgdesaix@colostate.edu/scratch/AMRE/all-samples/bamfiles/x2.0/",
                    replacement = "",
                    x = Bam))
amre.meta <- read_csv("../data/samples/amre_master_tidy.csv")
pops <- unique(amre.meta$K5_breeding) %>%
  unique() %>%
  sort()
```

Select only bams of interest for training (i.e. all breeding and no individuals from PA because their group is hard to resolve)

```{r}
nonbreeding.bams.meta <- amre.bams %>%
  left_join(amre.meta) %>%
  filter(Stage == "Nonbreeding") %>%
  select(full_path, Bam, Sample, Site)
```



### Checking assignment

```{r}

like_loo <- read_table("./nonbreeding-likelihood/amre.nonbreeding.ind148.ds_2x.sites-filter.top_50_each.pop_like.txt",
                       col_names = pops) 
# like_loo[like_loo == "-inf"] <- NA
# like_loo[like_loo == "nan"] <- NA

nonbreeding.assigned <- cbind(nonbreeding.bams.meta, like_loo)

cluster_colors <-  c(
  South = "#FF99FF", # pink/South
  SouthDakota = "#3399FF", # blue/SouthDakota
  Northeast = "#FFFF33", # yellow/Northeast
  Northwest = "#339933", # green/Northwest
  Newfoundland = "#CC0033" # red/Newfoundland
  # WMT = "#00ffff" # light blue
) 
```


```{r}
nonbreeding.summary <- nonbreeding.assigned %>%
  pivot_longer(cols = Newfoundland:SouthDakota,
               names_to = "AssignedPop",
               values_to = "AssignedLike") %>%
  group_by(Sample) %>%
  mutate(AssignedProb = round(exp(AssignedLike - min(AssignedLike)) / sum(exp(AssignedLike - min(AssignedLike))),2 )) %>%
  filter(AssignedLike == max(AssignedLike)) %>%
  ungroup() %>%
  select(-full_path)
# write_csv(x = nonbreeding.summary,
#           file = "./output/amre.nonbreeding.ind148.ds_2x.sites-filter.top_50_each.assignment_summary.csv")

nonbreeding.summary
```

Check number of accurate assignments

```{r}
nonbreeding.summary %>%
  filter(AssignedProb > 0.8) %>%
  nrow()
```



```{r}
site <- "BE"
nonbreeding.summary %>%
  filter(Site == site) %>%
  group_by(AssignedPop) %>%
  summarize(N = n()) %>%
  ungroup() %>%
  ggplot(aes(x="", y=N, fill = AssignedPop)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = cluster_colors) +
  theme_void() +
  theme(legend.position = "none")
```

```{r}
unique_sites <- unique(nonbreeding.summary$Site)
for(i in 1:length(unique_sites)){
  p.pie <- nonbreeding.summary %>%
  filter(Site == unique_sites[i]) %>%
  group_by(AssignedPop) %>%
  summarize(N = n()) %>%
  ungroup() %>%
  ggplot(aes(x="", y=N, fill = AssignedPop)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = cluster_colors) +
  theme_void() +
  theme(legend.position = "none")
  
  outname <- paste0("./plots/pie-assignment/p.", unique_sites[i], ".pie.png")
  ggsave(plot = p.pie,
         filename = outname,
         height = 4,
         width = 4,
         dpi = 150)
}
```






